---
title: "Data Challenge #1: Dairy Cow Visualizations"
author: "Vaishnavi Sathiyamoorthy"
format: html
editor: visual
---

## Set Up

```{r}
## load in the required libraries 
library(readr)
library(tidyverse)
library(here)
library(RColorBrewer)

## read in the data using here() function
state_milk_production <- read_csv(here("data/state_milk_production.csv"))
clean_cheese <- read_csv(here("data/clean_cheese.csv"))
fluid_milk_sales <- read_csv(here("data/fluid_milk_sales.csv"))
milk_products_facts <- read_csv(here("data/milk_products_facts.csv"))
milkcow_facts <- read_csv(here("data/milkcow_facts.csv"))

## disable scientific notation
options(scipen = 999)
```

## Figure 1

```{r}
# data points are colored 
ggplot(milkcow_facts, aes(avg_milk_cow_number, # x axis is num of milk cows
                          milk_production_lbs / 1000000, # y axis has milk production divided by 1 million
                          color = year)) + 
  geom_point() + # scatterplot
  labs(title = "The Average Number of Milk Cows vs Milk Production",
       x = "Average Number of Milk Cows",
       y = "Milk Production (lbs in Millions)",
       color = "Year")
```

*Figure 1. This scatterplot shows the average number of milk cows on the x axis and the milk production in millions on the y axis. The data points are colored by the year these data points were taken. As time has progressed it's seen that the number of milk cows has decreased and the milk production has increased. In fact, from 2000, the number of milk cows has stayed somewhat consistent, but the amount of milk production has increased by nearly 4e10 lbs.*

## Figure 2

```{r}
state_milk_production %>% 
  group_by(region, year) %>% # grouped by year and region
  summarize("total" = sum(milk_produced) / 1000000) %>% # total milk production is divided by 1 million
  ggplot(aes(x = year, # x axis is year
             y = total, # y axis is total consumption
             color = region)) + # colored by the regions
  geom_line() + # line graph
  facet_wrap(~region) + # different graph for each region
  labs(title = "Total Milk Production by Year by Region",
       x = "Year",
       y = "Total Milk Production (in millions)",
       color = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), # rotates the x axis tick titles by 45 degrees
        legend.position = "none")
```

*Figure 2. This is a line graph that shows how milk production has changed over the years based on the region. There seems to be an increase over time in the Lake States, Southern Plains, Pacific, and Mountains. There seems to be a decrease in the Appalachians and Delta States. Northeast and Northern Plains seems to remain relatively the same over time.*

## Figure 3

```{r}
fluid_milk_sales %>% 
  group_by(milk_type, year) %>% # grouped by milk type and then year
  filter(milk_type != "Total Production") %>% # the rows with milk type as total production are removed
  summarise("Total" = sum(pounds)) %>% # the total pounds is calculated based on how it is grouped
  ggplot(aes(x = year, # x axis is year
             y = log(Total), # y axis is the log of the total
             color = milk_type)) + # colored by milk type
  geom_line() + # line graph
  labs(title = "Milk Type Production by Year",
       x = "Year",
       y = "Log of Milk Production",
       color = "Type of Milk")
```

*Figure 3. This line plot shows the log of total milk production for each category by year. Over time, there has been an increase in eggnog and flavored non-whole milk production. The production of buttermilk, flavored whole milk, and whole milk went down. Skim milk production increased and then decreased over time.*

## Figure 4

```{r}
# new variable created
clean_cheese_long <- clean_cheese %>% 
  select(1:13) %>% # first 13 columns are chosen
  pivot_longer(cols = 2:13, # all columns except for year is selected for pivoting
               names_to = "Cheese_Category", 
               values_to = "Pounds")

clean_cheese_long %>%
  ggplot(aes(x = Year, # x axis is year
             y = Pounds, # y axis is in lbs
             fill = Cheese_Category)) + # fill of bar graph is based on cheese category
  geom_bar(stat = "identity", # bar plot. the values that are given are plotted 
           position = "stack") + # stacked bar plot
  labs(x = "Year", 
       y = "Cheese Consumption (lbs)", 
       title = "Cheese Category Pounds Over Time", 
       fill = "Cheese") +
  scale_fill_brewer(palette = "Set3")
```

*Figure 4. This is a bar graph that shows the amount of cheese consumed by each category, but also the overall cheese consumption. It is clear that over the years the amount of cheese consumed has increased. Mozzarella and Cheddar increased the most over this time period. Foods and Spreads has decreased over time. Swiss consumption has stayed consistent through the years.*

## Figure 5

```{r}
# new merged dataset
merged <- inner_join(milk_products_facts, # inner join of 2 datasets
                     milkcow_facts) %>% 
  mutate(total_ice_cream = frozen_ice_cream_regular + frozen_ice_cream_reduced_fat) # new column created

ggplot(merged, aes(x = year)) +
  geom_line(aes(y = total_ice_cream, # ice cream line
                color = "Total Ice Cream Consumption")) +
  geom_line(aes(y = fluid_yogurt, # yogurt line
                color = "Yogurt Consumption")) +
  scale_color_manual(values = c("Total Ice Cream Consumption" = "red", 
                                "Yogurt Consumption" = "green")) +
  labs(title = "Yogurt and Ice Cream Consumption Over Time",
       x = "Year", 
       y = "Consumption (lbs)", 
       color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

*Figure 5. This line graph looks at the ice cream consumption and yogurt consumption over time. The ice cream consumption is shown in red and the yogurt consumption is shown in green. Ice cream consumption has gradually decreased over time. Yogurt consumption has increase greatly.*
